---
title: "Homework 2"
subtitle: "Mahitha Jangeti"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1
```{r}
polsmonth_df = 
  read_csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(month = month(as.integer(month), label = TRUE, abbr = FALSE),
    year = as.integer(year),) %>% 
  mutate(president = if_else(prez_gop == 1, "gop", "dem")) %>% 
  select(-prez_gop, -prez_dem, -day)
  knitr::kable(head(polsmonth_df, 8))
```

**Cleaning snp data**
```{r}
snp_df =
  read_csv("./fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year")) %>% 
 mutate(
  month = month(as.integer(month), label = TRUE, abbr = FALSE),
  year  = as.integer(year),
  year  = if_else(year > 25, 1900 + year, 2000 + year))%>% 
  select(-day) %>% 
  relocate(year, month)
  knitr::kable(head(snp_df, 8))
```

**Cleaning unemployment data**
```{r}
unemployment_df = 
  read_csv("./fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "percentage_unemployment") %>% 
  mutate(    
    year = as.integer(year),
    month = recode(month,
                   jan = "January",
                   feb = "February",
                   mar = "March",
                   apr = "April",
                   may = "May",
                   jun = "June",
                   jul = "July",
                   aug = "August",
                   sep = "September",
                   oct = "October",
                   nov = "November",
                   dec = "December"))
  knitr::kable(head(unemployment_df, 8))
```


**Merging Datasets**
```{r}
final_df =
  left_join(polsmonth_df, snp_df, by = c("year", "month")) %>% 
  left_join(unemployment_df,by = c("year", "month")) %>% 
  relocate(year, month, president, close, percentage_unemployment)
  knitr::kable(head(final_df, 8))
```

## Five Thirty Eight Data Analysis

The polsmonth_df data set contains information on number of national politicians who are democratic vs republican. It included variables such as **gov_gop**, **sen_gop**, and **rep_gop** to indicate number of republican governers/senators on associated dates and **gov_dem** etc. to indicate democratice governers/senators on associated date. The **prez_gop** variable was mutated to represented if the president was republican through a binary format (later edited). 

The snp_df data set contains observations related to Standard & Poor's Stock market index, which is used as a representative measure of the stock market as a whole. This is indicated by the *close** closing values of the S&P stock index on the associated date (month and year).

The unemployment_df data set contains information on the percentage of unemployment throughout the different months of the year. This is indicated by **year** and **month** that was formatted through `pivot_longer`.

The final_df dataset has a total of `r nrow(final_df)` rows and `r ncol(final_df)` columns. It is merged through the year and month columns of all three datasets. The range of years goes from `r min(final_df$year)` to 
`r max(final_df$year)` and key variables include **close**, **percentage_unemployment** , **president**, and associated dates to see politics and its relation to outcomes in unemployment rates and values of the stock index. 



## Problem 2

```{r}
trashwheel_df = 
  read_excel("./fivethirtyeight_datasets/trashwheelcollection.xlsx", 
     sheet = "Mr. Trash Wheel",
     range = "A2:N709",
     na = c("NA",".","")) %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = as.integer(round(sports_balls)),
         year = as.integer(year),
         trash_wheel = "mr.trash wheel")
```

**Read and Clean Professor Trash Wheel and Gwynnda Trash Wheel sheet**
```{r}
professor_trashwheel_df =  
  read_excel( "./fivethirtyeight_datasets/trashwheelcollection.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:L134",
    na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  mutate(
    year = as.integer(year),
    trash_wheel = "professor trash wheel")
  


gwynns_trashwheel_df =  
  read_excel( "./fivethirtyeight_datasets/trashwheelcollection.xlsx",
    sheet = "Gwynns Falls Trash Wheel",
    range = "A2:K351",
    na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  mutate(
    year = as.integer(year),
    trash_wheel = "gwynns trash wheel")
```

**Combining Datasets**
```{r}
combinewheel_df = 
  bind_rows(trashwheel_df, professor_trashwheel_df, gwynns_trashwheel_df) %>% 
  janitor::clean_names() %>% 
  relocate(trash_wheel, dumpster) %>% 
  select(trash_wheel:homes_powered)
```

## Mr Trashwheel Data Analysis


## Problem 3

```{r}
zori_df = 
  read_csv("./zillow_data/zori.csv") %>% 
  janitor::clean_names() %>% 
  relocate(county_name) %>% 
  select(-state, county = county_name) %>% 
  mutate(county = county %>% 
            str_replace("New York County", "New York") %>% 
            str_replace("Kings County", "Kings") %>%
            str_replace("Queens County", "Queens") %>%
            str_replace("Bronx County", "Bronx")) %>%
  pivot_longer(
    x2015_01_31:x2024_08_31,
    names_to = "date",
    values_to = "zillow observed rent index")
  knitr::kable(head(zori_df, 8))


neighborhood_df = 
  read_csv("./zillow_data/zip_codes.csv") %>% 
  janitor::clean_names() %>% 
  relocate(county, neighborhood)
  knitr::kable(head(neighborhood_df, 8))
```

**Merging Datasets**

```{r}
zillow_df = 
  left_join(zori_df, neighborhood_df, by = c("county")) %>% 
  janitor::clean_names()
  knitr::kable(head(zillow_df, 8))
```


## Zillow Data Analysis

The dataset contains a total of `r nrow(zillow_df)` observations across `r ncol(zillow_df)` variables. 
There are `r n_distinct(zillow_df$zip_code)` unique ZIP codes represented and `r n_distinct(zillow_df$neighborhood)` unique neighborhoods.



